---
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cosmo
---
<style>
body {
    position: absolute;
    left: 0px;}
</style>

<!--html_preserve-->
<div class="container">
<!-- Static navbar -->
<nav class="navbar navbar-default">
  <div class="container-fluid">
    <div class="navbar-header">
      <button type="button" class="navbar-toggle collapsed" data-toggle="collapse" data-target="#navbar" aria-expanded="false" aria-controls="navbar">
        <span class="sr-only">Toggle navigation</span>
        <span class="icon-bar"></span>
        <span class="icon-bar"></span>
        <span class="icon-bar"></span>
      </button>
      <a class="navbar-brand" href="../index.html">Home</a>
    </div>
    <div id="navbar" class="navbar-collapse collapse">
      <ul class="nav navbar-nav">
        
        
          <li class="">
					<a href="/get_to_know_tek.html">Get To Know Tek</a>
	</li>              
				  
				  
	<li class="">
					<a href="/past_project.html">Past Project</a>
	</li>

	<li class="">
					<a href="/cheat_sheet.html">Cheat Sheet</a>
	</li>

      </ul>
    </div><!--/.nav-collapse -->
  </div><!--/.container-fluid -->
</nav><!--navbar navbar-default-->
</div>
<!--/html_preserve-->



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pkg=c("tm")
sapply(pkg, require, character=T)
```

<h1>Miscellaneous</h1><ol type="I">

  <h2><li>Updating R</li></h2><ol type="1">
  </ol><!--Updating R-->


  <h2><li>Loading R packages</li></h2><ol type="1">
There are multiple ways to load an R package. The most general way is to load the package using the `library` function or the `require` function on the script.

```{r, eval=F, comment=""}
library(some_package)
require(some_pacakge)
```

If one is using a GUI(Graphic User Interphase) such as the `RStudio`, then a simple click on the check box next to the installed package tab will load the desired library. 
However, what if one needs to load multiple packages at once? What if some of these packages are uninstalled? 
  
<h3><li>Loading Multiple R packages</li></h3>
A long list of required packages may make your script looking unclean. Moreover, when additional packages are needed, the `library` or `required` function is repeated again.
    
```{r, eval=F, comment=""}
library(package1)
library(package2)
library(package3)
library(package4)
library(package5)
library(package6)
```    
  
One can avoid messy script as above and use tidier looking script as below.
  
```{r, eval=F, comment=""}
pkg=c("package1", "package2", "package3", "package4", "package5", "package6")
sapply(pkg, require, character=T)
```

<h3><li>Automatically installing R packages if not installed</li></h3>
The code chunk below is a common way to install missing packages.
```{r, eval=F, comment=""}
if (!require(missing_packages)) install.packages(missing_packages)
```

For example, 
```{r, eval=F, comment=""}
missing_packages<-c("ggplot2", "tidyr")
if (!require(missing_packages)) install.packages(missing_packages)
```


Often times (especially sharing R scripts with co-workers who are novice R users), however, people forget to install R packages before running the script. 

In combination of the two listed items above (set hyperlink), The following code chunk is an elegant way to install and then load missing packages, if not previously installed.

```{r, eval=F, comment=""}
pkg=c("pkg1","pkg2", "pkg3", "missing_pkg1", "missing_pkg2")
missing_packages<- pkg[!(pkg %in% installed.packages()[,"Package"])]
if (!require(missing_packages)) install.packages(missing_packages)
sapply(pkg, require, character=T)
```

```{r, eval=F, comment=""}
pkg=c("ggplot2", "")
missing_packages<- pkg[!(pkg %in% installed.packages()[,"Package"])]
if (!require(missing_packages)) install.packages(missing_packages)
sapply(pkg, require, character=T)
```

  
  
  </ol><!--Loading R packages-->


  <h2><li>Source</li></h2><ol type="1">
  </ol><!---->




</ol><!--Miscellaneous-->