---
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: cosmo
---
<style>
body {
    position: absolute;
    left: 0px;}
</style>

<!--html_preserve-->
<div class="container">
<!-- Static navbar -->
<nav class="navbar navbar-default">
  <div class="container-fluid">
    <div class="navbar-header">
      <button type="button" class="navbar-toggle collapsed" data-toggle="collapse" data-target="#navbar" aria-expanded="false" aria-controls="navbar">
        <span class="sr-only">Toggle navigation</span>
        <span class="icon-bar"></span>
        <span class="icon-bar"></span>
        <span class="icon-bar"></span>
      </button>
      <a class="navbar-brand" href="../index.html">Home</a>
    </div>
    <div id="navbar" class="navbar-collapse collapse">
      <ul class="nav navbar-nav">
        
        
          <li class="">
					<a href="/get_to_know_tek.html">Get To Know Tek</a>
	</li>              
				  
				  
	<li class="">
					<a href="/past_project.html">Past Project</a>
	</li>

	<li class="">
					<a href="/cheat_sheet.html">Cheat Sheet</a>
	</li>

      </ul>
    </div><!--/.nav-collapse -->
  </div><!--/.container-fluid -->
</nav><!--navbar navbar-default-->
</div>
<!--/html_preserve-->



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

#Organizing Data<!--{-}-->

##Data Frame<!--{.tabset}-->

###1. Subsetting

####1) Column subset

a. The `$` operator <br>
*df$colmn name*

b. By the column number <br>
*df[,column number]*

c. By the column name <br>
*df[, \"column name\"]*

d. <a href="https://cran.r-project.org/web/packages/dplyr/dplyr.pdf">`dplyr` package </a> <br>
Utilize the function `select` from the `dplyr` package.<br>
*select(df, column name)* <br>
recommend using it in conjunction with the pipe operator `%>%` and other functions such as `filter`, `summarize` or `mutate`.

#####e.g.

```{r, comment=""}
head(mtcars)
```

```{r, eval=F, comment=""}
mtcars$mpg
mtcars[,1]
mtcars[,"mpg"]
mtcars %>%
  select(mpg)
```

```{r, echo=F, comment=""}
head(mtcars$mpg)
```
<br>

#####Note

* Subset multiple columns
```{r, eval=F, comment=""}
mtcars[,1:5]
mtcars[,c(1,3,5)]
mtcars[,c("mpg", "disp", "drat")]
```

* Complement
```{r, eval=F, comment=""}
mtcars[,-c(1,3,5)]

```

* S4 objects
Check `@` for S4. <br>
To be updated...



<br>

####2) Row subset

a. By the row number <br>
*df[row number,]*

b. By the row name <br>
*df[\"row name\"", ]*

#####e.g.

```{r, eval=F, comment=""}
mtcars[1,]
mtcars["Mazda RX4",]
```

```{r, echo=F, comment=""}
mtcars[1,]
```

###2. Filtering



###3. List

##dplyr

##tidyr