---
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: cosmo
---
<style>
body {
    position: absolute;
    left: 0px;}
</style>

<!--html_preserve-->
<div class="container">
<!-- Static navbar -->
<nav class="navbar navbar-default">
  <div class="container-fluid">
    <div class="navbar-header">
      <button type="button" class="navbar-toggle collapsed" data-toggle="collapse" data-target="#navbar" aria-expanded="false" aria-controls="navbar">
        <span class="sr-only">Toggle navigation</span>
        <span class="icon-bar"></span>
        <span class="icon-bar"></span>
        <span class="icon-bar"></span>
      </button>
      <a class="navbar-brand" href="../index.html">Home</a>
    </div>
    <div id="navbar" class="navbar-collapse collapse">
      <ul class="nav navbar-nav">
        
        
          <li class="">
					<a href="/get_to_know_tek.html">Get To Know Tek</a>
	</li>              
				  
				  
	<li class="">
					<a href="/past_project.html">Past Project</a>
	</li>

	<li class="">
					<a href="/cheat_sheet.html">Cheat Sheet</a>
	</li>

      </ul>
    </div><!--/.nav-collapse -->
  </div><!--/.container-fluid -->
</nav><!--navbar navbar-default-->
</div>
<!--/html_preserve-->



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

#Organizing Data<!--{-}-->

##I. List
List is the most general form of objects in `R`. <br>
`[` is used to access the sub-list, `[[` to access the element.

###1. Create List
a. `list()` function
```{r}
alnum=list(c("a","b","c","d"), c(1,2))
```

```{r, echo=F, comment=""}
alnum
```

The first sub-list of 'alnum' would be the character vector.
This character vector may be retrieved by running the following code.
```{r, comment=""}
alnum[[1]]
```
Similarly, the following code will retrieve the second element of the second sub-list.

```{r, comment=""}
alnum[[2]][2]
```

One can also name the sub-list simply adding labels before the sub list.

```{r}
alnum2=list(alphabet=c("e","f","g","h"), integer=c(3,4))
```

```{r, echo=F, comment=""}
alnum2
```

When the sub-lists are labeled, `[[` is no longer necessary.
```{r, comment=""}
alnum2$alphabet     #equivalent of alnum2[[1]]
alnum2$integer[2]   #equivalent of alnum2[[2]][2]
```

cf) `unlist()`

b. List as an output from a function
Functions such as `lapply` or `strsplit` generates its output in a list format.
```{r, comment=""}
lapply(1:2, function(x) c(x-1,x+1))
strsplit("abcdef", "c")
```

###2. Merge Multiple Lists
a. Into a list via `append` function
```{r, comment=""}
append(alnum, alnum2)
```

```{r, comment=""}
list(alnum, alnum2)
```
b. Into a Data Frame



##II. Data Frame<!--{.tabset}-->

###1. Create Data Frame

data.frame
as.data.frame
rbind.data.frame
cbind.data.frame

?as.data.frame()

Arguments:
stringsAsFactors: logical. `F` if the character vector should not be converted to a factor.
e.g.
```{r}
library(dplyr)
check=matrix(letters[1:9], nrow=3, byrow=T)
check %>% as.data.frame() %>% str()
check %>% as.data.frame(stringsAsFactors=F) %>% str()
check %>% as.data.frame() %>% .$V1
```



###2. Subsetting

####1) Column subset

a. The `$` operator <br>
*df$colmn name*

b. By the column number <br>
*df[,column number]*

c. By the column name <br>
*df[, \"column name\"]*

d. <a href="https://cran.r-project.org/web/packages/dplyr/dplyr.pdf">`dplyr` package </a> <br>
Utilize the function `select` from the `dplyr` package.<br>
*select(df, column name)* <br>
recommend using it in conjunction with the pipe operator `%>%` and other functions such as `filter`, `summarize` or `mutate`.

#####e.g.

```{r, comment=""}
head(mtcars)
```

```{r, eval=F, comment=""}
mtcars$mpg
mtcars[,1]
mtcars[,"mpg"]
mtcars %>%
  select(mpg)
```

```{r, echo=F, comment=""}
head(mtcars$mpg)
```
<br>

#####Note

* Subset multiple columns
```{r, eval=F, comment=""}
mtcars[,1:5]
mtcars[,c(1,3,5)]
mtcars[,c("mpg", "disp", "drat")]
```

* Complement
```{r, eval=F, comment=""}
mtcars[,-c(1,3,5)]

```

* S4 objects
Check `@` for S4. <br>
To be updated...



<br>

####2) Row subset

a. By the row number <br>
*df[row number,]*

b. By the row name <br>
*df[\"row name\"", ]*

#####e.g.

```{r, eval=F, comment=""}
mtcars[1,]
mtcars["Mazda RX4",]
```

```{r, echo=F, comment=""}
mtcars[1,]
```

###3. Filtering



###4. Data Presentation
Data set structured in a human friendly format dose not necessarily imply computer/machine friendly fashion or vise versa. Often times, therefore, data structure should be changed from long form to wide form or in the reversed manner. This may be even more important if one is sending and receiving data with co-workers who are not familiar with programming.  <a href="https://cran.r-project.org/web/packages/tidyr/tidyr.pdf">`tidyr`</a> and <a href="https://cran.rstudio.com/web/packages/reshape2/reshape2.pdf">`reshape`</a> are two useful packages that 


###5. Nested Data Frame


##tidyr