---
output:
  html_document:
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
```

# HTML Scraping and Parsing
Web scraping and parsing was(and will be) an important part of my research interest. There are many useful packages in `R` that makes web scraping simple: `rvest`, `XML`, `RCurl`, `curl`, `RSelenium`, and etc.

## Using the `rvest` package
```
install.packages("rvest") #if you haven't installed on yet
require(rvest) #load the rvest package
```

###`read_html`
read_html(*url*)

Argument
*url*: A url string

```{r, tidy=TRUE, comment=""}
read_html("https://jeffsong9.github.io/")
```

###`html_nodes`
* Usage: Parse `HTML` documents using XPath or CSS selector.
html\_nodes(*html\_doc*, *css*, *xpath*)

* Argument
*html\_doc*: A `HTML` document or a node set
*css*: css selector
*xpath*: xpath selector

* e.g.:
read_html(*some\_url*) %>%<br>
  html_nodes(xpath="//b")
Extracts all bold font text using the xpath selector //b.

read_html(*some\_url*) %>%<br>
  html_nodes(xpath='//div [@id="*something*"])

read_html(*some\_url*) %>%<br>
  html_nodes("center")
Extract using css selector center

* chain subsetting
read_html(*some\_url*) %>%<br>
  html_nodes("center") %>%<br>
  html_nodes("font")
  
```{r, comment=""}
read_html("https://jeffsong9.github.io/") %>%
  html_nodes(xpath="//div") %>%
  head(5)
```
  
  
cf) `html_node`

###`html_attr`
Usage: Extrac attributes with a given name
html_attr(*html\_doc*, *name*)

Arguments:
*html node*: A`HTML` document or a node set
*name*: Name of attribute to extract

e.g.
read_html(*some\_url*) %>%<br>
  html_nodes(xpath="//a") %>%<br>
  html_attr(name="href")

```{r, comment=""}
read_html("https://jeffsong9.github.io/") %>%
  html_nodes(xpath="//img [@class='photo']")
```

```{r, comment=""}
read_html("https://jeffsong9.github.io/") %>%
  html_nodes(xpath="//img [@class='photo']") %>%
  html_attr(name="src")
```

cf) html_attrs
