---
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: cosmo
---

<style>
body {
    position: absolute;
    left: 0px;}
</style>


<!--html_preserve-->
<div class="container">
<!-- Static navbar -->
<nav class="navbar navbar-default">
  <div class="container-fluid">
    <div class="navbar-header">
      <button type="button" class="navbar-toggle collapsed" data-toggle="collapse" data-target="#navbar" aria-expanded="false" aria-controls="navbar">
        <span class="sr-only">Toggle navigation</span>
        <span class="icon-bar"></span>
        <span class="icon-bar"></span>
        <span class="icon-bar"></span>
      </button>
      <a class="navbar-brand" href="../index.html">Home</a>
    </div>
    <div id="navbar" class="navbar-collapse collapse">
      <ul class="nav navbar-nav">
        
        
          <li class="">
					<a href="/get_to_know_tek.html">Get To Know Tek</a>
	</li>              
				  
				  
	<li class="">
					<a href="/past_project.html">Past Project</a>
	</li>

	<li class="">
					<a href="/cheat_sheet.html">Cheat Sheet</a>
	</li>

      </ul>
    </div><!--/.nav-collapse -->
  </div><!--/.container-fluid -->
</nav><!--navbar navbar-default-->
</div>
<!--/html_preserve-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
```

#HTML Scraping and Parsing
Web scraping and parsing was(and will be) an important part of my research interest. There are many useful packages in `R` that makes web scraping simple: `rvest`, `XML`, `RCurl`, `curl`, `RSelenium`, and etc. 

## Using the <a href="https://cran.r-project.org/web/packages/rvest/rvest.pdf">`rvest` package</a>
```{r, eval=F}
install.packages("rvest") #if you haven't installed `rvest` yet
require(rvest) #load the `rvest` package
```

###`read_html`
read_html(*url*)

Argument
*url*: A url string

```{r, tidy=TRUE, comment=""}
read_html("https://jeffsong9.github.io/")
```

###`html_nodes`
* Usage: Parse `HTML` documents using XPath or CSS selector.
html\_nodes(*html\_doc*, *css*, *xpath*)

* Argument
*html\_doc*: A `HTML` document or a node set
*css*: css selector
*xpath*: xpath selector

* e.g.:
read_html(*some\_url*) %>%<br>
  html_nodes(xpath="//b")
Extracts all bold font text using the xpath selector //b.

read_html(*some\_url*) %>%<br>
  html_nodes(xpath='//div [@id="*something*"])

read_html(*some\_url*) %>%<br>
  html_nodes("center")
Extract using css selector center

* chain subsetting
read_html(*some\_url*) %>%<br>
  html_nodes("center") %>%<br>
  html_nodes("font")
  
```{r, comment=""}
read_html("https://jeffsong9.github.io/") %>%
  html_nodes(xpath="//div") %>%
  head(5)
```
  
  
cf) `html_node`

###`html_attr`
Usage: Extrac attributes with a given name
html_attr(*html\_doc*, *name*)

Arguments:
*html node*: A`HTML` document or a node set
*name*: Name of attribute to extract

e.g.
read_html(*some\_url*) %>%<br>
  html_nodes(xpath="//a") %>%<br>
  html_attr(name="href")

```{r, comment=""}
read_html("https://jeffsong9.github.io/") %>%
  html_nodes(xpath="//img [@class='photo']")
```

```{r, comment=""}
read_html("https://jeffsong9.github.io/") %>%
  html_nodes(xpath="//img [@class='photo']") %>%
  html_attr(name="src")
```

cf) html_attrs




